Enzo: An Adaptive Mesh Refinement Code for Astrophysics

### 2013-06-08 overview.tex
Fixed some refs
Finished up my edits on the fluid equations.
More edits to fluid section - done?

\mqk{Move this table to the Appendix?}
\input{table_of_variables_and_constants}

@@

%\red{
%There is some debate over whether to put in the hydro equations without cosmology, in addition
%to the equations that are already here.  Would this
%be more or less clear?  We're leaning towards saying ``when cosmology is turned off,
% $\dot{a}$ goes to zero and $a$ goes to 1, and the poisson equation changes to blah blah blah'' 
%and assuming our readers can figure it out for themselves.
%We also want to emphasize that the heating and cooling is glossed over in this section, 
%and it's discussed more thoroughly later.
%} %end red
% GB: I agree this would be shorter, and maybe it is a better idea, but I think it's nice to be able
% GB: to include both equations and at the same time make the transformation to comoving coords
% GB: very clear -- this has engendered a lot of confusion in the past and one of the goals of writing
% GB this paper is to answer peoples questions before they ask them.

% GB: I commented out the following.  It has now been incorporated into the above discussion
% GB: (see my comments anove on why I think this is a good way to do it)
%
%\enzo\ solves the equations of ideal gas dynamics
%in a coordinate system that is comoving with the expanding universe:

%\begin{equation}
%\frac{\partial \rho_b}{\partial t} + \frac{1}{a} \vecv_b \cdot \grad \rho_b =  \label{enzoconserve}
% - \frac{1}{a} \rho_b \div \vecv_b 
%\end{equation}

%\begin{equation}
%\frac{\partial \vecv_b}{\partial t} + \frac{1}{a} (\vecv_b \cdot \grad) \vecv_b = \label{enzomomentum} 
%-\frac{\dot{a}}{a} \vecv_b - \frac{1}{a \rho_b} \grad p - \frac{1}{a} \grad \phi
%\end{equation}

%\begin{eqnarray}
%\frac{\partial E}{\partial t} + \frac{1}{a} \vecv_b \cdot \grad E =  
%- \frac{\dot{a}}{a} \left( 3 \frac{p}{\rho_b} + \vecv_b^2 \right)  \nonumber \\
%- \frac{1}{a \rho_b} \div (p \vecv_b) \nonumber \\
%- \frac{1}{a} \vecv_b \cdot \grad \phi + \Gamma - \Lambda
%\label{enzoenergy}
%\end{eqnarray}

%Where $\rho_b$ is the comoving baryon density, $\vecv_b$ is the baryon velocity, $p$ is the pressure, 
%$\phi$ is the modified gravitational potential (in comoving coordinates, which is related to
%the potential in proper coordinates $\Phi$ by $\phi \equiv \Phi + 0.5 $a\"{a}$ \vecx^2$) and 
%$a$ is the ``expansion parameter'' which describes the expansion of a smooth, homogeneous 
%universe as a function of time.  This expansion parameter is related to the redshift:  
%$a \equiv 1/(1+z)$.  All derivatives are in comoving coordinates.  $E$ is the specific
%energy of the gas (total energy per unit mass), and
%$\Gamma$ and $\Lambda$ represent radiative heating and cooling processes as described below.
%  Equations~\ref{enzoconserve}, \ref{enzomomentum}
%and \ref{enzoenergy} represent the conservation of mass, momentum and total (e.g.,
%kinetic plus thermal) fluid energy.

%The equations above are closed with three more equations:

%\begin{equation}
%E = p / [(\gamma - 1) \rho_b] + \vecv^2/2   \label{enzoeos}
%\end{equation}

%\begin{equation}
%\nabla^2 \phi = \frac{4 \pi G}{a} (\rho_b + \rho_{dm} - \rho_0 )
%\label{enzopoisson}
%\end{equation}

%\begin{equation}
%\frac{\ddot{a}}{a} = - \frac{4 \pi G}{3 a^3} (\rho_0 + 3 p_0 / c^2) + \Lambda/3.
%\label{enzocomove}
%\end{equation}

%where $\rho_{dm}$ is the comoving dark matter density, $\rho_0$ is the comoving
%background density 
%($\rho_0 \equiv \Omega_{matter} \rho_{crit}$) and $p_0$ is the background pressure, 
%$\gamma$ is the ratio of specific heats and $\Lambda$ is the cosmological constant.
%Equations~\ref{enzoeos}, \ref{enzopoisson} and~\ref{enzocomove} are the 
%equation of state, Poisson's equation in comoving form and an equation that 
%describes the evolution of the comoving coordinates (i.e. the formula for the 
%expansion of an isotropic, homogeneous universe). 
%All particles in the simulation are governed by Newton's equations in comoving 
%coordinates:

%\begin{equation}
%\frac{d \vecx_{part}}{dt} = \frac{1}{a} \vecv_{part}
%\label{enzopartvel}
%\end{equation}

%\begin{equation}
%\frac{d \vecv_{part}}{dt} = -\frac{\dot{a}}{a} \vecv_{part} - \frac{1}{a} \grad \phi
%\label{enzopartmom}
%\end{equation}

%Where $\vecx_{part}$ and $\vecv_{part}$ refer to the position and peculiar velocity of any
%particles in the system.  Note that the system of equations~\ref{enzoconserve}-\ref{enzopartmom} 
%is valid only in regimes where relativistic effects are not
%important ($v_b, v_{dm} \ll c$, where c is the speed of light) 
%and where cosmological curvature effects are 
%unimportant, meaning that the simulation volume is much smaller than the radius
%of curvature of the universe, as defined as $r_{hub} \equiv c/H_0$, where $c$ is the
%speed of light and $H_0$ is the Hubble constant.

%\dcc{Fixed punctuation, added 'no mhd, rhd' caveat to Zeus statements.  This has caused
%confusion in the past.}

%\subsubsection{Magnetohydrodynamics}

%\subsubsection{Radiation transport: ray-tracing}

@@

Four different (magneto)-hydrodynamic methods are implemented in
\enzo: (i) the hydrodynamic-only piecewise parabolic method (PPM)
developed by~\citet{1984JCoPh..54..174C} and extended to cosmology
by~\citet{1995CoPhC..89..149B}; (ii) the MUSCL-like Godunov scheme
\citep{MUSCL} with or without magnetic fields and Dedner-based
divergence cleaning, described in more detail in
\citet{WangAbelZhang08} and \citet{WangAbel09}; (iii) a constrained
transport staggered MHD scheme \citep{Collins10}, and (iv) the
second-order finite-difference hydrodynamics method described in
\zeus~\citep{Stone92a,Stone92b}.

@@

%The advantage of the PPM method is that it conserves energy better and is higher-order accurate.  The disadvantage is that it is generally less robust than the ZEUS method.

%%% If you prefer the following \paragraph{} sections to be numbered,
%%% then you can add \setcounter{secnumdepth}{5} to the preamble...

@@

\paragraph{MHD with Constrained Transport (MHD)}
This MHD method is second order in time and space, and preserves the
divergence constraint, $\div \vecB = 0$, to machine precision
through the Constrained Transport (CT) method.  CT, originally
described by \citet{Evans88}, updates the magnetic field with the curl
of an electric field, suitably formulated to preserve $\div
\vecB$.  We employ the CT methods described by \citet{Balsara99} and
\citet{Gardiner05} with the second order hyperbolic
solver of \citet{Li08a} and the divergence free AMR scheme of
\citet{Balsara01}.

>>>>

\input{table_of_variables_and_constants}

@@

@@

Four different (magneto)-hydrodynamic methods are implemented in
\enzo: (i) the hydrodynamic-only piecewise parabolic method (PPM)
developed by~\citet{1984JCoPh..54..174C} and extended to cosmology
by~\citet{1995CoPhC..89..149B}; (ii) the MUSCL-like Godunov scheme
\citep{1977JCoPh..23..276V} with or without magnetic fields and Dedner-based
divergence cleaning, described in more detail in
\citet{WangAbelZhang08} and \citet{WangAbel09}; (iii) a constrained
transport staggered MHD scheme \citep{Collins10}, and (iv) the
second-order finite-difference hydrodynamics method described in
\zeus~\citep{Stone92a,Stone92b}.

@@

@@

\paragraph{Godunov MHD with Constrained Transport (MHD)}
This MHD method is second order in time and space, and preserves the
divergence constraint, $\div \vecB = 0$, to machine precision
through the Constrained Transport (CT) method.  CT, originally
described by \citet{Evans88}, updates the magnetic field with the curl
of an electric field, suitably formulated to preserve $\div
\vecB$.  We employ the CT methods described by \citet{Balsara99} and
\citet{Gardiner05} with the second order hyperbolic
solver of \citet{Li08a} and the divergence free AMR scheme of
\citet{Balsara01}.
